### 2018-05-26 scp.tex
Remove some comments
Improve syntax and fix formating errors

Each $A_2$ is packed into a single continuous buffer.
Figure~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
Every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets, enabling
the inter-thread cache conflicts caused by $A_2$ to be
eliminated completely!

Figures~\ref{fig:partition.conventional} and~\ref{fig:partition.segmented}
essentially represent two different partitioning styles for a shared cache,
referred to as \emph{way-partitioning} and \emph{set-partitioning}, respectively.
To obtain the data layout in set-partitioning,
$A_2$ can no longer be stored in a single continuous buffer.
Instead, $A_2$ will be distributed to 12 continuous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While avoiding inter-thread cache conflicts,
set-partitioning
introduces some complexity to GEMM implementations
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continuous ones.

@@

Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
SCP will systematically determine the
the memory layout for the
packed matrices $A_2$ and $B_2$ used in GEMM.
The memory layout of the other packed matrices at
lower layers are determined by the memory layout of
$A_2$ and $B_2$.
We assume that the share matrix $B_2$ is packed
in the conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

@@

We use a memory descriptor $\mathcal{D}_t^M$ to specify 
the memory space occupied by a 
packed matrix $M$ used by thread $t$.
The memory descriptor is essentially a subspace of
the whole address space, i.e., $\mathcal{D}_t^M \subset \mathcal{A}$.
%% The memory layout of a packed matrix is described by a memory descriptor,
%% which is essentially a subspace of the whole address space $\mathcal{A}$.
%% Each packed matrix $M$ used by thread $t$,
%% is associated with a memory descriptor $\mathcal{D}_t^M$
%% specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses,
$S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partitioning, the memory descriptor contains only one segment.
With set-partitioning,
the memory descriptor contains a sequence of equally strided segments.

%% FIXME use l as the level iterator?
\begin{algorithm}
  \caption{SCP phase 1: allocate memory buffers}
  \label{alg:scp.phase1}
  \begin{algorithmic}[1]
    \REQUIRE $L_l = (c_l,l_l,n_l,nt_l)$ for $l \in [0, NL]$, $NT$,
    $M_r$, $N_r$, $K_c$, $M_c$, $N_c$
    \ENSURE $\mathcal{D}_p$ and $\mathcal{D}_s$
    \STATE $size_p \gets es M_c K_c$ \label{line:size.p}
    \STATE $size_s \gets es N_c K_c / NT$ \label{line:size.s}
    \STATE $align \gets l_1$ \label{line:align.init}
    \FOR {$l=1$ to $NL$} \label{line:align.for}
    \IF {$nt_l / nt_{l-1} > 1$ \AND $L_l$ is non-LRU} \label{line:align.type}
    \STATE $align \gets lcm(align, wc_l / nt_l)$ \label{line:align.update}
    \ENDIF
    \ENDFOR \label{line:align.endfor}
    \STATE $size_p \gets \lceil size_p / align \rceil \cdot align$ \label{line:align}
    %% \STATE $size \gets NT \cdot size_p + size_s$
    \STATE $addr_p \gets allocate(\mathcal{A}, size_p \cdot NT)$ \label{line:alloc.begin}
    \STATE $addr_s \gets allocate(\mathcal{A}, size_s \cdot NT)$ \label{line:alloc.end}
    \STATE $\mathcal{D}_p \gets \lbrace [addr_p, addr_p + size_p \cdot NT) \rbrace$ \label{line:d.begin}
    \STATE $\mathcal{D}_s \gets \lbrace [addr_s, addr_s + size_s \cdot NT) \rbrace$ \label{line:d.end}
  \end{algorithmic}
\end{algorithm}

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} describes
the second phase of SCP.
The main component is a recursive procedure $subspace$
(lines~\ref{line:subspace.begin} -- \ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierarchy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
This procedure
has three input parameters, $l$ and $idx$ for
identifying the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all the children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e.,
a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$,
causing $subspace$ to return early
(lines~\ref{line:shortcut.begin} -- \ref{line:shortcut.end}).
Otherwise, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(lines~\ref{line:for.begin} -- \ref{line:for.end}).
For each child node $cidx$ at layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branch (lines~\ref{line:if} -- \ref{line:endif})
represents two distinct ways for computing
$\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierarchy in a depth-first-search order.

>>>>

Each $A_2$ is packed into a single continuous buffer.
Figure.~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
Every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets, enabling
the inter-thread cache conflicts caused by $A_2$ to be
eliminated completely!

Figures~\ref{fig:partition.conventional} and~\ref{fig:partition.segmented}
essentially represent two different partitioning styles for a shared cache,
referred to as \emph{way-partitioning} and \emph{set-partitioning}, respectively.
To obtain the data layout in set-partitioning,
$A_2$ can no longer be stored in a single continuous buffer.
Instead, $A_2$ will be distributed to 12 continuous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
While avoiding inter-thread cache conflicts,
set-partitioning
introduces some complexity to GEMM implementations
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continuous ones.

@@

Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
SCP will systematically determine the
the memory layout for the packed matrices $A_2$ and $B_2$ used in GEMM.
The memory layout of the other packed matrices at
lower layers are determined once the memory layout of
$A_2$ and $B_2$ are determined.
We assume that the share matrix $B_2$ is packed
in the conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

@@

We use a memory descriptor $\mathcal{D}_t^M$ to specify 
the memory space occupied by a 
packed matrix $M$ used by thread $t$.
The memory descriptor is essentially a subspace of
the whole address space, i.e., $\mathcal{D}_t^M \subset \mathcal{A}$.
%% The memory layout of a packed matrix is described by a memory descriptor,
%% which is essentially a subspace of the whole address space $\mathcal{A}$.
%% Each packed matrix $M$ used by thread $t$,
%% is associated with a memory descriptor $\mathcal{D}_t^M$
%% specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
A segment is represented by its start and end addresses, $S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partitioning, the memory descriptor contains only one segment.
With set-partitioning,
the memory descriptor contains a sequence of equally strided segments.

\begin{algorithm}
  \caption{SCP phase 1: allocate memory buffers}
  \label{alg:scp.phase1}
  \begin{algorithmic}[1]
    \REQUIRE $L_l = (c_l,l_l,n_l,nt_l)$ for $l \in [0, NL]$, $NT$,
    $M_r$, $N_r$, $K_c$, $M_c$, $N_c$
    \ENSURE $\mathcal{D}_p$ and $\mathcal{D}_s$
    \STATE $size_p \gets es M_c K_c$ \label{line:size.p}
    \STATE $size_s \gets es N_c K_c / NT$ \label{line:size.s}
    \STATE $align \gets l_1$ \label{line:align.init}
    \FOR {$l=1$ to $NL$} \label{line:align.for}
    \IF {$nt_l / nt_{l-1} > 1$ \AND $L_l$ is non-LRU} \label{line:align.type}
    \STATE $align \gets lcm(align, wc_l / nt_l)$ \label{line:align.update}
    \ENDIF
    \ENDFOR \label{line:align.endfor}
    \STATE $size_p \gets \lceil size_p / align \rceil \cdot align$ \label{line:align}
    %% \STATE $size \gets NT \cdot size_p + size_s$
    \STATE $addr_p \gets allocate(\mathcal{A}, size_p \cdot NT)$ \label{line:alloc.begin}
    \STATE $addr_s \gets allocate(\mathcal{A}, size_s \cdot NT)$ \label{line:alloc.end}
    \STATE $\mathcal{D}_p \gets \lbrace [addr_p, addr_p + size_p \cdot NT) \rbrace$ \label{line:d.begin}
    \STATE $\mathcal{D}_s \gets \lbrace [addr_s, addr_s + size_s \cdot NT) \rbrace$ \label{line:d.end}
  \end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{alg:scp.phase2} describes the second phase of SCP.
The main component is a recursive procedure $subspace$
(lines~\ref{line:subspace.begin} -- \ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierarchy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
This procedure
has three input parameters, $l$ and $idx$ for
identifying the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all the children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e., a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$,
causing $subspace$ to return early
(lines~\ref{line:shortcut.begin} -- \ref{line:shortcut.end}).
Otherwise, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(lines~\ref{line:for.begin} -- \ref{line:for.end}).
For each child node $cidx$ at layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branch (lines~\ref{line:if} -- \ref{line:endif})
represents two distinct ways for computing
$\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierarchy in a depth-first-search order.
